---
title: "Project"
author: "Kameni"
date: "Sunday, December 21, 2014"
output: word_document
---
```{r}
Sys.setlocale(category = "LC_TIME", locale="English")

time <- format(Sys.time(),"%a, %b %d %Y, %X")


rversion <- R.Version()[[13]]

sysversion <- sessionInfo()[2]
```



```{r, echo=FALSE}
library(caret)
library(randomForest) 
library(rpart) 
library(rpart.plot)

# setting the overall seed for reproduceability
set.seed(1442)
```

####1 Data Loading & Explorating
First we want to load the data sets into R 
```{r}

trainingDf<-read.csv("./pml-training.csv")
dim(trainingDf)# Check number of variables and observations

testDf<-read.csv("./pml-testing.csv")
dim(testDf)# Check number of variables and observations


```


####2 Data Processing
in this part we face 2 problems:a) the missing values and b) the fact that the data set contains irrelevant variables.

#####a-) Processing missing values
we decide to Delete columns that contains missing values.

```{r}
trainingDf<-trainingDf[,colSums(is.na(trainingDf)) == 0];dim(trainingDf)
testDf <-testDf[,colSums(is.na(testDf)) == 0];dim(testDf)
```


#####b-) Processing irrelevant variables
As the first 7 columns are non-numeric non-integer, we decide to delete them.
```{r}
trainingDf <-trainingDf[,-c(1:7)];dim(trainingDf)
testDf <-testDf[,-c(1:7)];dim(testDf)
```


##### c) Adapting the columns of training and to testing data set

 We now make sure that both sets have the same number of columns.
 
```{r}
a<-0
colNames<-c()
nomTr<-names(trainingDf);nomTest<-names(testDf)
del<-setdiff( nomTest,trainingDf)
for(i in 1:length(nomTr)){
      if( nomTr[i] %in% nomTest){
          name<- nomTr[i]
          a<-a+1
          colNames[a]<-name
      }      
}
length(colNames)
colNames<-append(colNames,"classe");
trainingDf<-trainingDf[,colNames]
dim(trainingDf)

```


####3 Creating training and test 

In order to perform cross-validation, the training data set is partionned into 2 sets: subTraining (70%) and subTest (30%).
This will be performed using random subsampling without replacement.
```{r}
subsamples <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
subTraining <- trainingset[subsamples, ] 
subTesting <- trainingset[-subsamples, ]
dim(subTraining)
dim(subTesting)
head(subTraining)
head(subTesting)
```

